Fluent statistical computing interfaces for biological data analysis

### 2020-09-18 Rmd/05-embeddings.Rmd
fix screenshots layout, closes #6
fix typos

So when the the loss function is minimised this corresponds to large attractive
forces, that is, the pair-wise distances in $Y$ are small when there are
non-zero $p_{ij}$, i.e. $x_i$ and $x_j$ are close together. The repulsive force
should also be small when the loss function is minimised, that is, when
pair-wise distances in $Y$ are large regardless of the magnitude of the
corresponding distances in $X$.

@@

First, we downloaded the raw mouse retinal single cell RNA-seq data from
@Macosko2015-ot using the **scRNAseq** Bioconductor package [@scRNAseq-d]. We
have followed a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the the **scater**
by removing cells with high proportion of mitochondrial gene expression and low
numbers of genes detected, we log-transformed and normalised the expression 
values and  finally selected highly variable genes (HVGs) using **scran** 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighbours graph (with $k = 10$) and 
used Louvain clustering to generate clusters [@Blondel2008-bx].

@@

```{r mouse-01, fig.cap = "(ref:mouse-01)", out.height = "75%"}
display_screenshot("mouse-01", vid_lookup)
```

@@

From the video linked in figure \@ref(fig:mouse-01), we learn that the embedding has mostly captured the
clusters relative location to each other to their location in high dimensional space, with a notable exception of points in cluster 3 and 10 as shown with
linked brushing (figure \@ref(fig:mouse-01)a). As expected, t-SNE mitigates the crowding problem that is
an issue for tour in this case, where points in clusters 2,4,6, and 11 are
clumped together in tour view, but are blown up in the embedding view (figure \@ref(fig:mouse-01)b). 
The tour appears to form a tetrahedron like shape, with points lying on
surface and along the vertices of the the tetrahedron in 5-$d$ PCA space -
a phenomena that has also been observed in @Korem2015-af 
(figure \@ref(fig:mouse-01)c). Brushing on
the tour view, reveals points in cluster 9 that are diffuse in the embedding 
view, points in cluster 9 are relatively far away and spread apart from other 
clusters in the tour view, but has points placed in cluster 3 and 9 in the embedding
(figure \@ref(fig:mouse-01)d).

>>>>

So when the loss function is minimised this corresponds to large attractive
forces, that is, the pair-wise distances in $Y$ are small when there are
non-zero $p_{ij}$, i.e. $x_i$ and $x_j$ are close together. The repulsive force
should also be small when the loss function is minimised, that is, when
pair-wise distances in $Y$ are large regardless of the magnitude of the
corresponding distances in $X$.

@@

First, we downloaded the raw mouse retinal single cell RNA-seq data from
@Macosko2015-ot using the **scRNAseq** Bioconductor package [@scRNAseq-d]. We
have followed a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the **scater** package
by removing cells with high proportion of mitochondrial gene expression and low
numbers of genes detected, we log-transformed and normalised the expression 
values and  finally selected highly variable genes (HVGs) using **scran** 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighbours graph (with $k = 10$) and 
used Louvain clustering to generate clusters [@Blondel2008-bx].

@@

```{r mouse-01, fig.cap = "(ref:mouse-01)", out.height = "85%"}
display_screenshot("mouse-01", vid_lookup)
```

@@

From the video linked in figure \@ref(fig:mouse-01), we learn that the embedding has mostly captured the
clusters relative location to each other to their location in high dimensional space, with a notable exception of points in cluster 3 and 10 as shown with
linked brushing (figure \@ref(fig:mouse-01)a). As expected, t-SNE mitigates the crowding problem that is
an issue for tour in this case, where points in clusters 2,4,6, and 11 are
clumped together in tour view, but are blown up in the embedding view (figure \@ref(fig:mouse-01)b). 
The tour appears to form a tetrahedron like shape, with points lying on
surface and along the vertices of the tetrahedron in 5-$d$ PCA space -
a phenomena that has also been observed in @Korem2015-af 
(figure \@ref(fig:mouse-01)c). Brushing on
the tour view, reveals points in cluster 9 that are diffuse in the embedding 
view, points in cluster 9 are relatively far away and spread apart from other 
clusters in the tour view, but has points placed in cluster 3 and 9 in the embedding
(figure \@ref(fig:mouse-01)d).
